---
title: "QC Report"
output: 
  html_document:
    number_sections: true
    theme: sandstone
    highlight: espresso
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
# append_log <- function(..., Logfile="log_ParticulatorRMD.txt"){
#   try(
#     write(paste(Sys.time(), ...), file=Logfile, append = TRUE),
#     silent = TRUE
#   )
# }
# append_log("start")
# CUSTOMTEXT <- params$customtext
# if(CUSTOMTEXT=="") CUSTOMTEXT <- "100,300"

library(ggplot2)
library(data.table)
library(dplyr)
library(ggrepel)
library(plotly)
library(DT)
library(scales)
supertable <- function(df){
  datatable(df, extensions="Buttons", filter = list(position = 'top', clear = FALSE), 
            options=list(pageLength = nrow(df), dom="Bfrtip", buttons=c("copy","csv","excel","pdf","print")), rownames = F)
}

#set global variables
colorOK <- "deepskyblue"
colorMEH <- "orange"
colorBAD <- "red"

#read file
# read.csv("Syringe Pump Test 50 2022-08-11 13-55-48-ExperimentSummary.csv", sep=",", row.names=NULL, col.names = seq(100)) %>% 
list0 <- list.files(pattern = "csv$") %>% 
  lapply(function(x) {
    df <- read.csv(x, sep=",", row.names=NULL, col.names = seq(100)) %>% {.[,apply(.,2,function(x) !all(is.na(x)) )]}
    n <- ncol(df)-2
    measurements <- paste0("measure",seq(n-1))
    samplename <- gsub("ExperimentSummary.csv$","",x)
    
    sections <- c("[Experiment Details]","[Conditions]","[Settings]","[Results]",
                  "[Concentration Warnings]","[Information]","[Size Data]","Bin centre (nm)","Percentile")
    section_lengths <- c(8,12,7,8,7,9,10,1000,101) %>% `names<-`(sections)
    section_table <- df[,1] %>% {which(. %in% names(section_lengths))} %>% 
      {data.frame(row.names=names(section_lengths), start=.+1, end=.+as.vector(section_lengths))}
    
    get_rows <- function(i, n_cols=n, transpose=T){
      tab <- section_table[sections[i],,drop=T] %>% unlist() %>% {seq(.[1],.[2])} %>% {df[.,,drop=F]}
      if(transpose) tab <- tab %>% {`rownames<-`(.[, 2:n_cols,drop=F], tab[,1])} %>% t() %>% as.data.frame() %>% 
          mutate_if(~sum(as.numeric(.x))>0, ~as.numeric(.x))
      return(tab)
    }
    
    details <- get_rows(1, n_cols=2) %>% mutate(sample=samplename)
    general <- lapply(2:3, get_rows) %>% do.call(cbind, .) %>% mutate(sample=samplename)
    results <- get_rows(4) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    warnings <- get_rows(5) %>% mutate(sample=samplename)
    info <- get_rows(6) %>% mutate(sample=samplename)
    size_data <- get_rows(7) %>% mutate(sample=samplename)
    
    graph_data1 <- get_rows(8, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("size_bin",measurements)) %>% 
      melt(id.vars="size_bin", measure.vars=measurements) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    graph_data2 <- get_rows(9, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("percentile",measurements)) %>% 
      melt(id.vars="percentile", measure.vars=measurements) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    
    return(list(details=details, general=general, results=results, warnings=warnings, info=info, size_data=size_data,
                graph_data1=graph_data1, graph_data2=graph_data2))
  })

details <- list0 %>% lapply(function(x) x$details) %>% do.call(rbind, .)
general <- list0 %>% lapply(function(x) x$general) %>% do.call(rbind, .)
results <- list0 %>% lapply(function(x) x$results) %>% do.call(rbind, .)
warnings<- list0 %>% lapply(function(x) x$warnings)%>% do.call(rbind, .)
info    <- list0 %>% lapply(function(x) x$info)    %>% do.call(rbind, .)
size_data<- list0 %>%lapply(function(x) x$size_data)%>% do.call(rbind, .)
graph_data1 <- list0 %>% lapply(function(x) x$graph_data1) %>% do.call(rbind, .)
graph_data2 <- list0 %>% lapply(function(x) x$graph_data2) %>% do.call(rbind, .)

graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df)))
  
  ggplot(df, aes_string(xmetric2, "sample")) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    stat_summary(aes(label=round(..x..,2)) , fun=mean, geom="label") +
    labs(x=xmetric) +
    theme_bw() +
    theme(axis.title.y=element_blank())
}

graph2d <- function(df, xmetric, ymetric){
  ggplot(df, aes_string(xmetric, ymetric, color="sample")) +
    geom_point(alpha=.4, size=3) +
    theme_bw() +
    theme()
}

graph3d <- function(df, xmetric, ymetric, labelmetric){
  graph2d(df, xmetric=xmetric, ymetric=ymetric) +
    geom_text_repel(aes_string(label=labelmetric))
}
```

# General Information {.tabset}

- creation date of this file: `r format(Sys.time(), '%B %d, %Y')`
- operator: `r unique(details[,"Operator Name"])`
- camera: `r unique(general[,"Camera Type"])`
- laser: `r unique(general[,"Laser Type"])`
- software: `r unique(details[,"Software Version"])`

## Sample Table

```{r}
# append_log("sample table1")
supertable(details[,c(4,2,5)] %>% `colnames<-`(gsub("\\."," ",colnames(.))))
```

## Conditions

```{r}
supertable(general[,c(1,2,5:12)] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

## Settings

```{r results="asis"}
supertable(general[,13:19] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

# Results {.tabset}

## Table

```{r}
sections_table <- colnames(results) %>% {.[grepl("^sample$|Concentration Upgrade|Dilution factor", .)]} %>% {.[c(3,2,1)]}
supertable(results[,sections_table] %>% `colnames<-`(gsub("\\(.*","",colnames(.))))
```

```{r results="asis"}
graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df)))
  
  ggplot(df, aes_string(xmetric2, "sample")) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    stat_summary(aes(label=round(..x..,2)) , fun=median, geom="label") +
    labs(x=xmetric) +
    theme_bw() +
    theme(axis.title.y=element_blank())
}

sections_graphs <- colnames(results) %>% {.[!. %in% sections_table]}
for(section in sections_graphs){
  
  cat("\n") 
  cat("##", gsub("\\([^\\)]*\\)","", gsub(" per ","/", section)), "\n") # Create second level headings
  
  print(graph1d(results, section))
  
  cat("\n")
  
}
```

# Warnings & Info {.tabset}

```{r}
graphHeat <- function(df){
  n_records <- length(unique(df$sample))
  n_samples <- length(unique(gsub("\\(.*","",df$sample)))
  records_per_sample <- n_records/n_samples
  hlines <- seq(records_per_sample,n_records-records_per_sample,by=records_per_sample) + 0.5
  ggplot(df, aes(variable, sample, fill=fillcolor)) +
    geom_tile(color="white") +
    geom_text(aes(label=value)) +
    geom_hline(yintercept=hlines, size=1) +
    scale_fill_identity() +
    theme_bw() +
    theme(axis.title=element_blank(), axis.text.x=element_text(angle=45, vjust=1, hjust=1))
}
```


## Warnings

```{r}
toplot <- warnings %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")")) %>% melt(id.vars="sample") %>%
  mutate(fillcolor=ifelse(value %in% c("No","OK",""), "lightblue", "red"), value=gsub(" ","\n",value))

graphHeat(toplot)
```

## Info

```{r}
toplot <- info %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")"),
                          Concentration=gsub(" concentration","",Concentration)) %>% 
  melt(id.vars="sample") %>% 
  mutate(fillcolor=ifelse(value %in% c("No","OK",""), "lightblue", "red"), value=gsub(" ","\n",value))

graphHeat(toplot)
```

# Size Data {.tabset}

## Mean, Mode, SD

```{r}
ggplot(size_data, aes(Mean, Mode)) +
  geom_point() +
  facet_wrap(~sample) +
  geom_rug(color="red") +
  geom_errorbar(aes(xmin=Mean-SD, xmax=Mean+SD)) +
  geom_errorbar(aes(ymin=Mode-SD, ymax=Mode+SD)) +
  scale_x_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
  scale_y_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, hjust=0, vjust=.5))

```

## D10, D50, D90

```{r}
toplot <- melt(size_data, id.vars="sample", measure.vars=c("D10","D50","D90"), value.name="value")
ggplot(toplot, aes(value, variable, color=sample)) +
  geom_boxplot() + 
  geom_point(alpha=.4) +
  geom_hline(yintercept=seq(2)+.5, color="gray") +
  scale_x_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
  theme_bw() +
  theme(panel.grid.major.y=element_blank(), axis.title.y=element_blank())

```

## valid tracks

```{r}
graph1d(size_data, "Valid Tracks")
```

# Detailed Graphs {.tabset}

## Size Percentiles

```{r}
ggplot(graph_data2, aes(percentile, value, color=factor(variable))) +
  geom_path() +
  facet_wrap(~sample, scales="free_y") +
  geom_rug(aes(x=NULL), color="gray30", alpha=.2) +
  labs(y="particle size (nm)", x="percentile", color="measurement") +
  scale_y_continuous(expand=c(0,0), limits=c(0,1000), breaks=seq(0,1000,by=100)) +
  scale_x_continuous(expand=c(0,0), limits=c(0,100)) +
  theme_bw() +
  theme()
```

## Size Distribution

```{r}
ggplot(graph_data1, aes(size_bin, value, color=factor(variable))) +
  geom_path() +
  facet_wrap(~sample) +
  scale_y_continuous(label=comma, limits=c(0,max(graph_data1$value)*1.1), expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,1000,by=200), limits=c(0,1000)) +
  labs(x="particle size (nm)", y="concentration (particles/mL)", color="measurement") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))
```

