---
title: "QC Report"
output: 
  html_document:
    number_sections: true
    theme: sandstone
    highlight: espresso
params:
  customtext: "100,300"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
LOGFILE <- paste0("log_particulatorRMD.txt")
append_log <- function(..., Logfile=LOGFILE){
  try(
    write(paste(Sys.time(), ...), file=Logfile, append = TRUE),
    silent = TRUE
  )
}
do <- function(anything, Logfile=LOGFILE){
  out <- try(anything)
  if(class(out) %in% "try-error") append_log(out, Logfile=Logfile)
  return(out)
}
append_log("start")
customtext <- params$customtext
append_log(customtext)
if(customtext==""){
  CUSTOMTEXT <- c(100,300)
}else{
  CUSTOMTEXT <- as.numeric(strsplit(customtext,",")[[1]])
}
append_log(CUSTOMTEXT)
MINVAL <- CUSTOMTEXT[1]
MAXVAL <- CUSTOMTEXT[2]

append_log("load libraries")
do({
  library(ggplot2)
  library(data.table)
  library(dplyr)
  library(ggrepel)
  library(plotly)
  library(DT)
  library(scales)
})
supertable <- function(df, extrafilter=F){
  myfilter <- list()
  if(extrafilter) myfilter <- list(position = 'top', clear = FALSE)
  datatable(df, extensions="Buttons", filter = myfilter, 
            options=list(pageLength = nrow(df), dom="Bfrtip", buttons=c("copy","csv","excel","pdf","print")), rownames = F)
}

#set global variables
colorOK <- "deepskyblue"
colorMEH <- "orange"
colorBAD <- "red"

#read file
# read.csv("Syringe Pump Test 50 2022-08-11 13-55-48-ExperimentSummary.csv", sep=",", row.names=NULL, col.names = seq(100)) %>% 
append_log("list")
list0 <- list.files(pattern = "ExperimentSummary.csv$") %>% {.[1]} %>% 
  lapply(function(x) {
    df <- read.csv(x, sep=",", row.names=NULL, col.names = seq(100)) %>% {.[,apply(.,2,function(x) !all(is.na(x)) )]}
    n <- ncol(df)-2
    measurements <- paste0("measure",seq(n-1))
    samplename <- gsub("ExperimentSummary.csv$","",x)
    
    sections <- c("[Experiment Details]","[Conditions]","[Settings]","[Results]",
                  "[Concentration Warnings]","[Information]","[Size Data]","Bin centre (nm)","Percentile")
    section_starts <- df[,1] %>% {which(. %in% sections)}
    section_lengths <- c(8,12,7,8,7,9,10,section_starts[9]-section_starts[8]-1,101) %>% `names<-`(sections)
    section_table <- df[,1] %>% {which(. %in% names(section_lengths))} %>% 
      {data.frame(row.names=names(section_lengths), start=.+1, end=.+as.vector(section_lengths))}
    
    get_rows <- function(i, n_cols=n, transpose=T){
      tab <- section_table[sections[i],,drop=T] %>% unlist() %>% {seq(.[1],.[2])} %>% {df[.,,drop=F]}
      if(transpose) tab <- tab %>% {`rownames<-`(.[, 2:n_cols,drop=F], tab[,1])} %>% t() %>% as.data.frame() %>% 
          mutate_if(~sum(as.numeric(.x))>0, ~as.numeric(.x))
      return(tab)
    }
    
    details <- get_rows(1, n_cols=2) %>% mutate(sample=samplename)
    general <- lapply(2:3, get_rows) %>% do.call(cbind, .) %>% mutate(sample=samplename)
    results <- get_rows(4) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    warnings <- get_rows(5) %>% mutate(sample=samplename)
    info <- get_rows(6) %>% mutate(sample=samplename)
    size_data <- get_rows(7) %>% mutate(sample=samplename)
    
    graph_data1 <- get_rows(8, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("size_bin",measurements)) %>% 
      melt(id.vars="size_bin", measure.vars=measurements) %>% mutate(sample=samplename) %>%  mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    graph_data2 <- get_rows(9, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("percentile",measurements)) %>% 
      melt(id.vars="percentile", measure.vars=measurements) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    
    return(list(details=details, general=general, results=results, warnings=warnings, info=info, size_data=size_data,
                graph_data1=graph_data1, graph_data2=graph_data2))
  })

append_log("extract data")
details <- list0 %>% lapply(function(x) x$details) %>% do.call(rbind, .)
general <- list0 %>% lapply(function(x) x$general) %>% do.call(rbind, .)
results <- list0 %>% lapply(function(x) x$results) %>% do.call(rbind, .)
warnings<- list0 %>% lapply(function(x) x$warnings)%>% do.call(rbind, .)
info    <- list0 %>% lapply(function(x) x$info)    %>% do.call(rbind, .)
size_data<- list0 %>%lapply(function(x) x$size_data)%>% do.call(rbind, .)
graph_data1 <- list0 %>% lapply(function(x) x$graph_data1) %>% do.call(rbind, .) %>% mutate(variable=as.character(variable))
graph_data2 <- list0 %>% lapply(function(x) x$graph_data2) %>% do.call(rbind, .) %>% mutate(variable=as.character(variable))

append_log("define functions")
graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df)))
  
  ggplot(df, aes_string(xmetric2, "sample")) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    stat_summary(aes(label=round(..x..,2)) , fun=mean, geom="label") +
    labs(x=xmetric) +
    theme_bw() +
    theme(axis.title.y=element_blank())
}

graph2d <- function(df, xmetric, ymetric){
  ggplot(df, aes_string(xmetric, ymetric, color="sample")) +
    geom_point(alpha=.4, size=3) +
    theme_bw() +
    theme()
}

graph3d <- function(df, xmetric, ymetric, labelmetric){
  graph2d(df, xmetric=xmetric, ymetric=ymetric) +
    geom_text_repel(aes_string(label=labelmetric))
}
```

# General Information {.tabset}

- creation date of this file: `r format(Sys.time(), '%B %d, %Y')`
- operator: `r unique(details[,"Operator Name"])`
- camera: `r unique(general[,"Camera Type"])`
- laser: `r unique(general[,"Laser Type"])`
- software: `r unique(details[,"Software Version"])`

## Sample Table

```{r}
append_log("table1")
supertable(details[,c(4,2,5)] %>% `colnames<-`(gsub("\\."," ",colnames(.))))
```

## Conditions

```{r}
append_log("table2")
supertable(general[,c(1,2,5:12)] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

## Settings

```{r results="asis"}
append_log("table3")
supertable(general[,13:19] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

# Detailed Graphs {.tabset}

## Size Distribution

The black line indicates the average of all measurements for the respective sample.

```{r}
append_log("Size Distribution")
avg <- group_by(graph_data1, size_bin, sample) %>% summarise(variable="avg", value=mean(value, na.rm=T))
ggplot(graph_data1, aes(size_bin, value, color=factor(variable))) +
  geom_rect(aes(xmin=MINVAL, xmax=MAXVAL, ymin=-Inf, ymax=Inf), 
            inherit.aes=F, data=data.frame(), fill="gray", alpha=.15, color=NA) +
  geom_path() +
  geom_path(data=avg, color="black", size=1) +
  facet_wrap(~sample) +
  scale_y_continuous(label=comma, limits=c(0,max(graph_data1$value)*1.1), expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,1000,by=200), limits=c(0,1000)) +
  labs(x="particle size (nm)", y="concentration (particles/mL)", color="measurement") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))
```

## Particles in Size Range

Note that the concentration is given in: k cells/mL = 1000 cells/mL

```{r}
append_log("Particles in size range")
do({
  myrange <- paste0(MINVAL,"-",MAXVAL,"nm")
  toplot <- graph_data1 %>% mutate(size=ifelse(size_bin<=MAXVAL & size_bin>=MINVAL, myrange, "other")) %>% 
    group_by(variable, sample, size) %>% summarise(area=sum(value))
  toplot2 <- merge(subset(toplot, size==myrange), subset(toplot, size=="other"), by=c("variable", "sample")) %>% dplyr::rename(area="area.x")
  mycolor <- "blue"
  
  ggplot(toplot, aes(variable, area/1000, fill=size)) +
    facet_wrap(~sample) +
    geom_bar(stat="identity", position="dodge") +
    geom_label(aes(label=paste0(round(area/1000),"\n",round(area/(area+area.y)*100),"%") ), data=toplot2, fill="white", color=mycolor, alpha=.8) +
    scale_y_continuous(limits=c(0,max(toplot$area, na.rm=T)*1.1/1000), expand=c(0,0)) +
    scale_fill_manual(values=`names<-`(c(mycolor, "gray"), c(myrange,"other"))) +
    labs(x="measurement", y="area under the curve (k particles/mL)") +
    theme_bw() +
    theme()
})
```


## Size Percentiles

The black line indicates the average of all measurements for the respective sample.

```{r}
append_log("Size Percentiles")
do({
  avg <- group_by(graph_data2, percentile, sample) %>% summarise(variable="avg", value=mean(value, na.rm=T))
  ggplot(graph_data2, aes(percentile, value, color=variable)) +
    geom_rect(aes(ymin=MINVAL, ymax=MAXVAL, xmin=-Inf, xmax=Inf), inherit.aes=F, data=data.frame(), fill="gray", alpha=.15, color=NA) +
    geom_path() +
    geom_path(data=avg, color="black", size=1) +
    facet_wrap(~sample, scales="free_y") +
    geom_rug(aes(x=NULL), color="gray30", alpha=.2) +
    labs(y="particle size (nm)", x="percentile", color="measurement") +
    scale_y_continuous(expand=c(0,0), limits=c(0,1000), breaks=seq(0,1000,by=100)) +
    scale_x_continuous(expand=c(0,0), limits=c(0,100)) +
    theme_bw() +
    theme()
})
```


# Warnings & Info {.tabset}

```{r}
append_log("Warnings and Info")
graphHeat <- function(df){
  n_records <- length(unique(df$sample))
  n_samples <- length(unique(gsub("\\(.*","",df$sample)))
  records_per_sample <- n_records/n_samples
  p <- ggplot(df, aes(variable, sample, fill=fillcolor)) +
    geom_tile(color="white") +
    geom_text(aes(label=value)) +
    scale_fill_identity() +
    theme_bw() +
    theme(axis.title=element_blank(), axis.text.x=element_text(angle=45, vjust=1, hjust=1))
  append_log("... plot generated")
  if(n_samples>1){
    hlines <- seq(records_per_sample,n_records-records_per_sample,by=records_per_sample) + 0.5
    p <- p + geom_hline(yintercept=hlines, size=1)
  }
  append_log("... return plot")
  return(p)
}
```

## Warnings

```{r}
append_log("Warnings")
do({
  toplot <- warnings %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")")) %>% melt(id.vars="sample") %>%
    mutate(fillcolor=ifelse(value %in% c("No","OK",""), "lightblue", "yellow"), value=gsub(" ","\n",value))
  append_log("... table generated")
  # graphHeat(toplot)
})
```

## Info

```{r}
append_log("Info")
do({
  toplot <- info %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")"),
                            Concentration=gsub(" concentration","",Concentration)) %>% 
    melt(id.vars="sample") %>% 
    mutate(fillcolor=ifelse(value %in% c("No","OK",""), "lightblue", "yellow"), value=gsub(" ","\n",value))
  
  graphHeat(toplot)
})
```


# Results {.tabset}

## Table

```{r}
append_log("Results - Table")
do({
  supertable(results %>% dplyr::relocate(sample) %>% `colnames<-`(gsub("\\(.*","",colnames(.))))
})
```

```{r results="asis"}
graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df)))
  
  ggplot(df, aes_string(xmetric2, "sample")) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    stat_summary(aes(label=round(..x..,2)) , fun=median, geom="label") +
    labs(x=xmetric) +
    theme_bw() +
    theme(axis.title.y=element_blank())
}

sections_graphs <- colnames(results) %>% {.[!. %in% colnames(results)[1:3]]}
for(section in sections_graphs){
  
  cat("\n") 
  cat("##", gsub("\\([^\\)]*\\)","", gsub(" per ","/", section)), "\n") # Create second level headings
  
  print(graph1d(results, section))
  
  cat("\n")
  
}
```

# Size Data {.tabset}

## Mean, Mode, SD

```{r}
append_log("Size Data - Mean Mode SD")
do({
  ggplot(size_data, aes(Mean, Mode)) +
    geom_point() +
    facet_wrap(~sample) +
    geom_rug(color="red") +
    geom_errorbar(aes(xmin=Mean-SD, xmax=Mean+SD)) +
    geom_errorbar(aes(ymin=Mode-SD, ymax=Mode+SD)) +
    scale_x_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
    scale_y_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=.5))
})
```

## D10, D50, D90

```{r}
append_log("Size Data - D10 D50 D90")
do({
  toplot <- melt(size_data, id.vars="sample", measure.vars=c("D10","D50","D90"), value.name="value")
  ggplot(toplot, aes(value, variable, color=sample)) +
    geom_boxplot() + 
    geom_point(alpha=.4) +
    geom_hline(yintercept=seq(2)+.5, color="gray") +
    scale_x_continuous(limits=c(0,1000), expand=c(0,0), breaks=seq(0,1000,200)) +
    theme_bw() +
    theme(panel.grid.major.y=element_blank(), axis.title.y=element_blank())
})
```

## valid tracks

```{r}
append_log("Size Data - valid tracks")
do({
  graph1d(size_data, "Valid Tracks")
})
```


