---
title: "QC Report"
output: 
  html_document:
    number_sections: true
    theme: sandstone
    highlight: espresso
params:
  customtext: "100,300"
---














```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
#script explanation: 
"
generate a QC report from partile analyzer results
TARGET FOLDER: contains file(s) ending with ExperimentSummary.csv
CUSTOM TEXT: upper and lower particle size, separated by comma (e.g. 100,300). If not specified, uses default values
"
#=======================================================
#=======================================================
#=======================================================
LOGFILE <- "logfile_scriptSelector.txt"
PROGRESSFILE <- "scriptSelector_running.txt"
CUSTOM_TEXT <- params$customtext

append_log <- function(..., Logfile=LOGFILE){
  try(
    write(paste(Sys.time(), ...), file=Logfile, append = TRUE),
    silent = TRUE
  )
}
do <- function(anything, Logfile=LOGFILE){
  out <- try(anything)
  if(class(out) %in% "try-error") append_log(out, Logfile=Logfile)
  return(out)
}

append_log("_", Logfile=PROGRESSFILE)
append_log("Particulator", CUSTOMTEXT)
#=======================================================
#=======================================================
#=======================================================
if(CUSTOM_TEXT==""){
  customtext <- c(100,300)
}else{
  customtext <- as.numeric(strsplit(CUSTOM_TEXT,",")[[1]])
}
MINVAL <- customtext[1]
MAXVAL <- customtext[2]

append_log("load packages")
library(ggplot2)
library(data.table)
library(dplyr)
library(ggrepel)
library(plotly)
library(DT)
library(scales)
bignum <- function(x){
  format(as.numeric(x), scientific=F, big.mark=",")
}
supertable <- function(df, extrafilter=F){
  myfilter <- list()
  if(extrafilter) myfilter <- list(position = 'top', clear = FALSE)
  datatable(df, extensions="Buttons", filter = myfilter, 
            options=list(pageLength = nrow(df), dom="Bfrtip", buttons=c("copy","csv","excel","pdf","print")), rownames = F)
}

#set global variables
colorOK <- "deepskyblue"
colorMEH <- "orange"
colorBAD <- "red"
MAX_PARTICLE_SIZE <- 600

#read file
# read.csv("Syringe Pump Test 50 2022-08-11 13-55-48-ExperimentSummary.csv", sep=",", row.names=NULL, col.names = seq(100)) %>% 
append_log("list")
list0 <- list.files(pattern = "ExperimentSummary.*.csv$") %>% 
  lapply(function(x) {
    df <- read.csv(x, sep=",", row.names=NULL, col.names = seq(100)) %>% {.[,apply(.,2,function(x) !all(is.na(x)) )]}
    n <- ncol(df)-2
    measurements <- paste0("measure",seq(n-1))
    samplename <- gsub(" 202.-..-...*$","",x)
    
    sections <- c("[Experiment Details]","[Conditions]","[Settings]","[Results]",
                  "[Concentration Warnings]","[Information]","[Size Data]","Bin centre (nm)","Percentile")
    section_starts <- df[,1] %>% {which(. %in% sections)}
    section_lengths <- c(8,12,7,8,7,9,10,section_starts[9]-section_starts[8]-1,101) %>% `names<-`(sections)
    section_table <- df[,1] %>% {which(. %in% names(section_lengths))} %>% 
      {data.frame(row.names=names(section_lengths), start=.+1, end=.+as.vector(section_lengths))}
    
    get_rows <- function(i, n_cols=n, transpose=T){
      tab <- section_table[sections[i],,drop=T] %>% unlist() %>% {seq(.[1],.[2])} %>% {df[.,,drop=F]}
      if(transpose) tab <- tab %>% {`rownames<-`(.[, 2:n_cols,drop=F], tab[,1])} %>% t() %>% as.data.frame() %>% 
          mutate_if(~sum(as.numeric(.x))>0, ~as.numeric(.x))
      return(tab)
    }
    
    details <- get_rows(1, n_cols=2) %>% mutate(sample=samplename)
    general <- lapply(2:3, get_rows) %>% do.call(cbind, .) %>% mutate(sample=samplename)
    results <- get_rows(4) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    warnings <- get_rows(5) %>% mutate(sample=samplename)
    info <- get_rows(6) %>% mutate(sample=samplename)
    size_data <- get_rows(7) %>% mutate(sample=samplename)
    
    graph_data1 <- get_rows(8, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("size_bin",measurements)) %>% 
      melt(id.vars="size_bin", measure.vars=measurements) %>% mutate(sample=samplename) %>%  mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    graph_data2 <- get_rows(9, transpose=F) %>% {.[1:n]} %>% `colnames<-`(c("percentile",measurements)) %>% 
      melt(id.vars="percentile", measure.vars=measurements) %>% mutate(sample=samplename) %>% mutate_if(~!any(is.na(as.numeric(.x))), ~as.numeric(.x))
    
    return(list(details=details, general=general, results=results, warnings=warnings, info=info, size_data=size_data,
                graph_data1=graph_data1, graph_data2=graph_data2))
  })

append_log("extract data")
details <- list0 %>% lapply(function(x) x$details) %>% do.call(rbind, .)
general <- list0 %>% lapply(function(x) x$general) %>% do.call(rbind, .)
results <- list0 %>% lapply(function(x) x$results) %>% do.call(rbind, .)
warnings<- list0 %>% lapply(function(x) x$warnings)%>% do.call(rbind, .)
info    <- list0 %>% lapply(function(x) x$info)    %>% do.call(rbind, .)
size_data<- list0 %>%lapply(function(x) x$size_data)%>% do.call(rbind, .)
graph_data1 <- list0 %>% lapply(function(x) x$graph_data1) %>% do.call(rbind, .) %>% mutate(variable=as.character(variable))
graph_data2 <- list0 %>% lapply(function(x) x$graph_data2) %>% do.call(rbind, .) %>% mutate(variable=as.character(variable))

append_log("define functions")
graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df)))
  
  ggplot(df, aes_string(xmetric2, "sample")) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    stat_summary(aes(label=round(..x..,2)) , fun=mean, geom="label") +
    labs(x=xmetric) +
    theme_bw() +
    theme(axis.title.y=element_blank())
}

graph2d <- function(df, xmetric, ymetric){
  ggplot(df, aes_string(xmetric, ymetric, color="sample")) +
    geom_point(alpha=.4, size=3) +
    theme_bw() +
    theme()
}

graph3d <- function(df, xmetric, ymetric, labelmetric){
  graph2d(df, xmetric=xmetric, ymetric=ymetric) +
    geom_text_repel(aes_string(label=labelmetric))
}
```

# General Information {.tabset}

- creation date of this file: `r format(Sys.time(), '%B %d, %Y')`
- operator: `r unique(details[,"Operator Name"])`
- camera: `r unique(general[,"Camera Type"])`
- laser: `r unique(general[,"Laser Type"])`
- software: `r unique(details[,"Software Version"])`

## Sample Table

```{r}
append_log("table1")
supertable(details[,c(9,4,5,2)] %>% `colnames<-`(gsub("\\."," ",colnames(.))))
```

## Conditions

```{r}
append_log("table2")
supertable(general[,c(20,1,2,5:12)] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

## Settings

```{r results="asis"}
append_log("table3")
supertable(general[,c(20,13:19)] %>% `colnames<-`(gsub("\\/","\\/ ",colnames(.))))
```

# Detailed Graphs {.tabset}

## Size Distribution {.tabset}

The colored lines are the averages of all measurements for that sample. The black line represents the average of all averages of the samples.

```{r}
size_distribution_graph <- function(df){
  dfavg <- group_by(df, size_bin) %>% summarise(value=mean(value,na.rm=T), sample="avg")
  p <- ggplot(df, aes(size_bin, value, color=sample)) +
    geom_rect(aes(xmin=MINVAL, xmax=MAXVAL, ymin=-Inf, ymax=Inf),
              inherit.aes=F, data=data.frame(), fill="gray", alpha=.15, color=NA) + #does not work in plotly
    geom_path() +
    geom_path(data=dfavg, color="black")  +
    scale_y_continuous(limits=c(0,max(df$value)*1.1), expand=c(0,0), labels=function(x) format(x, scientifc=T)) +
    scale_x_continuous(expand=c(0,0), breaks=seq(0,1000,by=200), limits=c(0,MAX_PARTICLE_SIZE)) +
    labs(x="particle size (nm)", y="concentration (particles/mL)", color="average of all measurements") +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))
  return(ggplotly(p))
}
```

### Average

```{r}
append_log("Size Distribution average")
df <- group_by(graph_data1, size_bin, sample=gsub(" ","\n",sample)) %>% summarise(value=mean(value, na.rm=T))
size_distribution_graph(df)
```

### Median

```{r}
append_log("Size Distribution median")
df <- group_by(graph_data1, size_bin, sample=gsub(" ","\n",sample)) %>% summarise(value=median(value, na.rm=T))
size_distribution_graph(df)
```

## Particles in Size Range

Note that the concentration is given in: k cells/mL = 1000 cells/mL

```{r}
append_log("Particles in size range")
myrange <- paste0(MINVAL,"-",MAXVAL,"nm")
toplot <- graph_data1 %>% 
  mutate(size=ifelse(size_bin<=MAXVAL & size_bin>=MINVAL, myrange, "other")) %>% 
  group_by(variable, sample, size) %>% 
  summarise(area=sum(value))
mycolor <- "gold"

p <- ggplot(toplot, aes(size, area, fill=size)) +
  geom_boxplot(outlier.shape = NA, position="dodge") +
  geom_point(shape=21, size=3) +
  facet_wrap(~sample) +
  scale_y_continuous(limits=c(0,max(toplot$area, na.rm=T)*1.1), expand=c(0,0), labels=function(x) format(x, scientifc=T) ) +
  scale_fill_manual(values=`names<-`(c(mycolor, "gray"), c(myrange,"other"))) +
  labs(x="measurement", y="area under the curve (particles/mL)") +
  theme_bw() +
  theme(axis.title.x=element_blank())

ggplotly(p)
```

## Size Levels {.tabset}

```{r}
colorcode <- c("0-100"="yellow", "100-200"="limegreen", ">200"="lightblue")
toplot <- graph_data1 %>% 
  mutate(size=ifelse(size_bin>200 , names(colorcode)[1], ifelse(size_bin>=100, names(colorcode)[2], names(colorcode)[3])),
         size=factor(size, levels=rev(names(colorcode))),
         sample=gsub(" |_","\n",sample)) %>% 
  group_by(sample, size) %>% 
  summarise(particles=sum(value)) %>% 
  group_by(sample) %>% 
  arrange(sample, desc(size)) %>% 
  mutate(percentage=round(100*particles/sum(particles, na.rm=T),1),
         stacked_particles=cumsum(particles),
         stacked_percentage=cumsum(percentage))
```

### percentages

```{r}
p <- ggplot(toplot, aes(sample, percentage)) +
  geom_bar(aes(fill=size), stat="identity") +
  geom_text(aes(y=stacked_percentage-percentage/2, label=round(percentage,1) )) +
  # geom_text(aes(y=0, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[1]), vjust=0, show.legend = F) +
  # geom_text(aes(y=50, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[2]), vjust=.5, show.legend = F) +
  # geom_text(aes(y=100, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[3]), vjust=1, show.legend = F) +
  labs(fill="size (nm)") +
  scale_fill_manual(values=colorcode) +
  labs(x="measurement", y="area under the curve %") +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.text.x=element_text(angle=45, hjust=0, vjust=1))

ggplotly(p)
```

### absolute values

```{r}
p <- ggplot(toplot, aes(sample, particles)) +
  geom_bar(aes(fill=size), stat="identity") +
  geom_text(aes(y=stacked_particles-particles/2, label=format(particles, scientific=T,digits=2) )) +
  # geom_text(aes(y=5, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[1]), vjust=0, show.legend = F) +
  # geom_text(aes(y=50, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[2]), vjust=.5, show.legend = F) +
  # geom_text(aes(y=95, label=round(percentage,1)), data=subset(toplot, size==names(colorcode)[3]), vjust=1, show.legend = F) +
  labs(fill="size (nm)", y="particles/ml") +
  scale_fill_manual(values=colorcode) +
  labs(x="measurement", y="area under the curve (particles/mL)") +
  theme_bw() +
  theme(axis.title.x=element_blank())

ggplotly(p)
```



## Size Percentiles {.tabset}

The black line indicates the average of all measurements for the respective sample.

```{r}
size_percentiles_plot <- function(df){
  p <- ggplot(df, aes(percentile, value, color=sample)) +
    geom_rect(aes(ymin=MINVAL, ymax=MAXVAL, xmin=-Inf, xmax=Inf), inherit.aes=F, data=data.frame(), fill="gray", alpha=.15, color=NA) +
    geom_path() +
    # facet_wrap(~sample, scales="free_y") +
    geom_rug(aes(x=NULL), color="gray30", alpha=.2) +
    labs(y="particle size (nm)", x="percentile", color="measurement") +
    scale_y_continuous(expand=c(0,0), limits=c(0,MAX_PARTICLE_SIZE), breaks=seq(0,1000,by=100)) +
    scale_x_continuous(expand=c(0,0), limits=c(0,100)) +
    theme_bw() +
    theme()
  return(ggplotly(p))
}
```

### average

```{r}
avg <- group_by(graph_data2, percentile, sample) %>% summarise(value=mean(value, na.rm=T))
size_percentiles_plot(avg)
```

### median

```{r}
med <- group_by(graph_data2, percentile, sample) %>% summarise(value=median(value, na.rm=T))
size_percentiles_plot(med)
```


# Warnings & Info {.tabset}

```{r}
graphHeat <- function(df){
  n_records <- length(unique(df$sample))
  n_samples <- length(unique(gsub("\\(.*","",df$sample)))
  records_per_sample <- n_records/n_samples
  p <- ggplot(df, aes(variable, sample, fill=info)) +
    geom_tile(color="white") +
    scale_fill_manual(values=c(OK="lightblue", No="lightblue"), na.value="yellow") +
    theme_bw() +
    theme(axis.title=element_blank(), axis.text.x=element_text(angle=45, vjust=1, hjust=1), legend.position="none")
  append_log("... plot generated")
  if(n_samples>1){
    hlines <- seq(records_per_sample,n_records-records_per_sample,by=records_per_sample) + 0.5
    p <- p + geom_hline(yintercept=hlines, size=1)
  }
  append_log("... return plot")
  return(ggplotly(p))
}
```

## Warnings

```{r}
append_log("Warnings")
do({
  toplot <- warnings %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")")) %>% melt(id.vars="sample") %>%
    dplyr::rename(info="value")
  append_log("... table generated")
  graphHeat(toplot)
})
```

## Info

```{r}
append_log("Info")
do({
  toplot <- info %>% mutate(sample=paste0(sample,"(", seq(nrow(.)), ")"),
                            Concentration=gsub(" concentration","",Concentration)) %>% 
    melt(id.vars="sample") %>% 
    dplyr::rename(info="value")
  
  graphHeat(toplot)
})
```


# Results {.tabset}

## Table

```{r}
append_log("Results - Table")
do({
  df <- results
  df$`Concentration (Particles / ml)` <- bignum(df$`Concentration (Particles / ml)`)
  supertable(df %>% dplyr::relocate(sample))
})
```

```{r results="asis"}
graph1d <- function(df, xmetric){
  xmetric2 <- make.names(xmetric)
  df <- `colnames<-`(df, make.names(colnames(df))) %>% 
    mutate(sample=gsub(" ","\n",sample))
  
  p <- ggplot(df, aes_string("sample", xmetric2)) +
    geom_boxplot(outlier.shape=NA) +
    geom_point(alpha=.4) +
    # stat_summary(aes(label=round(..x..,2)) , fun=median, geom="label") +
    labs(y=xmetric) +
    theme_bw() +
    theme(axis.title.x=element_blank())
  ggplotly(p)
}

sections <- colnames(results) %>% {.[!. %in% colnames(results)[c(1,2,3,ncol(results))]]}
i <- 1
```

## `r sections[i]`

```{r}
graph1d(results, sections[i])
i <- i+1
```

## `r sections[i]`

```{r}
graph1d(results, sections[i])
i <- i+1
```

## `r sections[i]`

```{r}
graph1d(results, sections[i])
i <- i+1
```

## `r sections[i]`

```{r}
graph1d(results, sections[i])
i <- i+1
```

## `r sections[i]`

```{r}
graph1d(results, sections[i])
```


# Size Data {.tabset}

## Mean, Mode, SD

Mean and Mode are the x and y axis, respecively. The standard deviation (SD) is indicated by the size of the cross. Each point has a cross that spans 2x the SD in both x and y direction. The mode (a.k.a. modal value) is the most frequently occurring value (presumably after the data was binned).


```{r}
append_log("Size Data - Mean Mode SD")
do({
  p <- ggplot(size_data, aes(Mean, Mode)) +
    geom_point() +
    facet_wrap(~sample) +
    geom_rug(color="red") +
    geom_errorbar(aes(xmin=Mean-SD, xmax=Mean+SD)) +
    geom_errorbar(aes(ymin=Mode-SD, ymax=Mode+SD)) +
    scale_x_continuous(limits=c(0,MAX_PARTICLE_SIZE), expand=c(0,0), breaks=seq(0,1000,100), name="mean particle size (nm)") +
    scale_y_continuous(limits=c(0,MAX_PARTICLE_SIZE), expand=c(0,0), breaks=seq(0,1000,100), name="mode particle size (nm)") +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=.5))
  ggplotly(p)
})
```

## D10, D50, D90

```{r}
append_log("Size Data - D10 D50 D90")
do({
  toplot <- melt(size_data, id.vars="sample", measure.vars=c("D10","D50","D90"), value.name="value")
  p <- ggplot(toplot, aes(variable, value)) +
    geom_boxplot() + 
    facet_wrap(~sample) +
    geom_point(alpha=.4) +
    geom_hline(yintercept=seq(2)+.5, color="gray") +
    scale_y_continuous(limits=c(0,MAX_PARTICLE_SIZE), expand=c(0,0), breaks=seq(0,1000,200), name="particle size (nm)") +
    theme_bw() +
    theme(panel.grid.major.x=element_blank(), axis.title.x=element_blank())
  ggplotly(p)
})
```

## valid tracks

```{r}
append_log("Size Data - valid tracks")
do({
  graph1d(size_data, "Valid Tracks")
})
```

```{r}
unlink(PROGRESSFILE)
```

